Using Hybrid Machine Learning for Fundamental Stellar Parameter Inference

### 2024-06-10 paper/paper2/paper.tex
Updates from Overleaf

\begin{abstract}
    The advent of machine learning (ML) methods has and continues to revolutionize numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference.
    ML methods in spectroscopy are expected to perform with greater efficacy than conventional computational techniques.
    In this study we introduce a hybrid ML (HML) method that combines automatic differentiation (autodiff), interpolation, and Bayesian optimization to infer stellar parameters given stellar spectra.
    The stellar parameters we study are $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method has potential for extension to other properties such as [$\alpha$/Fe], C/O, and $f_{sed}$.
    We first use the semi-empirical approach to treating spectra as sets of tunable spectral lines introduced in \texttt{blase} \citep{blase}, contrasting with traditional data-driven approaches. 
    \texttt{blase} is used on 1,314 spectra from the PHOENIX synthetic spectral model grid \citep{PHOENIX}, identifying 128,723 unique spectral lines.
    For each of these lines, a manifold mapping stellar parameters to spectral line parameters is created using regular grid linear interpolation.
    These manifolds are aggregated to to create the PHOENIX generator, sporting parallelized reconstruction of spectra at continuously valued points within the PHOENIX subset.
    Gaussian Process minimization is then used to infer stellar parameters by minimizing the root-mean-square (RMS) loss between the input spectrum and PHOENIX generator spectra.
    From testing, the inference error in $T_{\mathrm{eff}}$ was 185 K, $\log(g)$ was 0.19, and [Fe/H] was 0.12 dex.
    Our products are threefold: first an archive of the state dictionaries of the \texttt{blase} clones of the PHOENIX subset, second the PHOENIX generator itself, and third, the reconstruction and inference algorithm.
    This study represents a rudimentary proof of concept showing that semi-empirical HML is a viable, albeit fledgling alternative to traditional approaches.
\end{abstract}

\section{Introduction}
Stellar spectra are exceedingly rich sources of information about the stars that produce them. 
Spectra encode fundamental properties such as temperature, surface gravity, and chemical composition via their seemingly countless absorption lines. 
Extrinsic properties such as the stellar radial velocity or the projected equatorial rotation further transform spectra. 
Our observations of the spectra degrade the otherwise pure astronomical signals, and so the resolution the observing instrument and other instrumental properties limit our ability to extract fundamental properties perfectly. 
Stellar spectra represent extremely complex, data, influenced by multiple parameters, that have gone through multiple transformations before reaching our detectors.

The challenge lies in extracting as much information as possible from a stellar spectrum to get back its stellar parameters. 
Spectroscopic surveys such as APOGEE develop their own in-house pipelines to extract stellar parameters from spectra (ASPCAP in the case of APOGEE), and these pipelines appear effective \citep{APOGEE, ASPCAP}. 
However, they all work on the core assumption that the data to be analyzed is a list of pixels. 
Analysis pipelines may be closed-source or limited to the scope of the survey itself, causing a sort of siloing effect among surveys. 
This motivates the development of more universal, instrument-agnostic open-source frameworks that can apply broadly to a range of spectral observations with relatively little tuning.

Multiple efforts have been made in this direction, treating spectra in different ways. 
The standard practice is to treat the wavelength and flux as simply two arrays and use bespoke algorithms tailored to a small number of well-calibrated spectral lines to obtain fundamental stellar parameters and chemical compositions. 
Other whole-spectrum fitting abstractions decompose model spectra into an eigenbasis, implementing the data-model comparison stage as a tractable regression, such as \texttt{starfish} \citep{starfish}. 
However, \texttt{blase} takes an intermediate approach \citep{blase}, treating spectra not as a set of pixels or a set of eigenbasis coefficients but as a set of spectral lines, specifically Voigt profiles.
Each of these approaches has tradeoffs, but one key distinction places \texttt{blase} in another category defined by the ability to fit the nonlinear spectral line parameters with autodiff.
We see autodiff used in tools such as \texttt{exojax} and \texttt{wobble} \citep{exojax, wobble}, but the recasting of spectra into sets of inherently nonlinear spectral lines positions \texttt{blase} as a unique and promising semi-empirical tool.
The original \texttt{blase} paper demonstrated the ability to tune spectral lines with autodiff, but only on a few hand-picked fiducial models, not scaled out to an entire model grid.

Ideally this process could be monolithic, with the end-to-end spectral inference code powered by a single autodiffable machine learning framework, like PyTorch or JAX \citep{pytorch, jax}. 
However, here we separate the problem into three pieces. 
First, the scaling out of the blase method to 1,314 precomputed synthetic spectral model clones, yielding a downloadable archive of pretrained machine learning models with 128,723 unique spectral lines. 
Second, the creation of manifolds mapping stellar parameters to spectral line parameters using regular grid linear interpolation. 
Finally, we show how reconstructions of the spectra using said manifolds can be used for spectral inference.
An overview of this process is shown in \autoref{fig:figure1}.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure1}
    \caption{Overview of the process used in this study.}
    \label{fig:figure1}
\end{figure*}
Advancements to this underlying basis have the potential to be top tier options with further research and development.

@@

\section{Cloning the PHOENIX Model Grid}
\subsection{The PHOENIX Subset}
For the purposes of this study, we did not consider the full range of the PHOENIX synthetic spectral model grid. 
Instead, we focused on a subset of the grid, with details given in \autoref{tab:table1}.

\begin{table}[h!]
    \centering
    \begin{tabular}{lll}
        \hline
        \bf{Parameter} & \bf{Symbol} & \bf{Range}\\
        \hline
        Alpha Element Abundance & $\alpha$ & [0] dex\\
        Iron Abundance & [Fe/H] & [-0.5, 0] dex\\
        Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K\\
        Surface Gravity & $\log(g)$ & [2, 6]\\
        Wavelength & $\lambda$ & [8038, 12849] \AA\\
        \hline
    \end{tabular}
    \caption{The subset of the PHOENIX grid used in this study. 
    These limits were imposed to reduce the computational cost of the algorithms and to ensure a rectilinear parameter space in order to work with \texttt{scipy}'s \texttt{RegularGridInterpolator} \citep{scipy}. 
    The wavelength limits in particular roughly line up with that of the Habitable Zone Planet Finder (HPF) spectrograph \citep{HPF}. 
    This subset is comprised of 1,314 individual spectra: 73 unique $T_{\mathrm{eff}}$ values, 9 unique $\log(g)$ values, and 2 unique [Fe/H] values.}
    \label{tab:table1}
\end{table}

\subsection{Preprocessing with \texttt{gollum}}
First, the PHOENIX subset was programmatically retrieved with \texttt{gollum} \citep{gollum}. The spectra were then put through a 
three-step preprocessing pipeline similar to that from \citealt{blase}.
\begin{enumerate}
    \item \textit{Blackbody Division}: Since the $T_{\mathrm{eff}}$ of each spectrum is known, the according blackbody spectrum was divided out.
    \item \textit{Percentile Normalization}: The spectra were normalized by dividing them by their 99th percentile.
    \item \textit{Continuum Normalization}: The spectra were further normalized by dividing them by a \nth{5} order polynomial continuum.
\end{enumerate}
Mathematically, we can express the preprocessing as follows:
\begin{gather}
    \mathsf{\bar{S}} = \dfrac{\mathsf{S}}{\mathsf{B}\mathsf{Q}_5P_{99}}
\end{gather}
where $\mathsf{\bar{S}}$ is the preprocessed spectrum, $\mathsf{S}$ is the original spectrum,  $\mathsf{B}$ is the blackbody spectrum, $\mathsf{Q}_n$ is the $n^\mathrm{th}$ order polynomial continuum fit, and $P_n$ is the $n^\mathrm{th}$ percentile function. 
Arithmetic operations between arrays are assumed to be elementwise in all notation from here on out.

\subsection{Line Identification with \texttt{blase}}
The next step was to convert the PHOENIX subset into a physically interpretable intermediate representation: a table of spectral line properties rather than a list of fluxes. 
This was done using \texttt{blase}, which detects spectral lines as Voigt profiles and tunes the profiles to mimic the original PHOENIX spectrum with back propagation. 
Four parameters were optimized: the line center $\mu$, the log-amplitude $\ln(a)$, the Gaussian width $\sigma$, and the Lorentzian width $\gamma$. 
The optimization used the Adam optimizer with a learning rate of 0.05 over 100 epochs \citep{adam}. 
In addition, we limited two custom parameters: wing cut to 6000 and prominence to 0.005. 
Wing cut is a parameter that determines the extent of the Voigt profile to evaluate, saving computational resources by not evaluating negligible line wings. 
Prominence sets a lower limit for the amplitude of detected lines, which also saves resources by disregarding shallow lines. 
In short, smaller values for wing cut and smaller values for prominence both decrease the computational cost of \texttt{blase}'s cloning process at the expense of decreasing its accuracy slightly. 
In addition, it should be mentioned that \texttt{blase} uses the pseudo-Voigt approximation, which saves on computational cost while remaining accurate to about 1\% \citep{pseudovoigt1, pseudovoigt2}. 
It is a weighted average of a Gaussian and Lorentzian as opposed to a convolution.
\texttt{blase}'s pseudo-Voight profile implementation uses the following:
\begin{subequations}
\begin{gather}
    \tilde{\mathsf{V}}_\mu(\lambda) = a\left[\eta \mathbf{L}(\lambda - \mu'; f) + (1 - \eta)\mathbf{G}(\lambda - \mu'; f)\right]\\
    \eta = \sum_{n=1}^3 \mathbf{u}_n\left(\dfrac{2\gamma}{f}\right)^n\\
    f = 32\sum_{n=0}^5 \mathbf{v}_n \left(\sqrt{2\ln(2)}\sigma\right)^{5-n}(\gamma)^n\\
    \nonumber \mathbf{u} = \mqty[1.36603 \\ -0.47719 \\ 0.11116] \quad \mathbf{v} = \mqty[1 \\ 2.69269 \\ 2.42843 \\ 4.47163 \\ 0.07842 \\ 1]
    \label{eq:equation2}
\end{gather}
\end{subequations}
where $\mathbf{L}$ and $\mathbf{G}$ are abbreviations for Lorentzian and Gaussian profiles, respectively. 
Notice that we use $\mu'$ instead of $\mu$ in the formula. 
This is because \texttt{blase} optionally allows the line center to shift slightly during optimization, and it is this shifted center which is used in computation. 
The individual Voigt profiles are still indexed by $\mu$ for cross-model line identification, explained in the next section. 
Once optimization was complete, the list of identified lines, was saved to a `state dictionary': a common representation for pre-trained machine learning models that can be stored to disk for reuse later. 
These are stored in the \texttt{.pt} file format for each of the 1,314 PHOENIX subset grid points. 
The total disk space these files take up is 465 MB (382 MB when downloaded as a zip archive).

@@

There is also a second issue. Spectral lines were often only detected in some spectra from the PHOENIX subset. 
In \autoref{fig:figure2}, we show that different grid points sport differing counts of detected spectral lines.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure2}
    \caption{Number of detected spectral lines at each grid point of 
    a slice of the PHOENIX subset at solar metallicity. We can see that
    the number of detected lines decreases with increasing $T_{\mathrm{eff}}$.
    Also note that from $T_{\mathrm{eff}} = 7000$ K onward, the spacing 
    between grid points increases from 100 K to 200 K.}
    \label{fig:figure2}
\end{figure*}

This missing-lines phenomenon can arise for three different reasons: Stellar atmospherse genuinely do not produce that line at detectable strength at the given temperature and surface gravity. 
Second, PHOENIX missed it, possibly accidentally. 
Or third, our line-finding and line-association algorithms missed it. 
Whatever the cause, these missing lines have immediate practical consequences. 
Rectilinear interpolation schemes break in regions where a line does not appear; you can't interpolate a quantity that simply doesn't exist.

@@

\section{Bayesian Inference and Testing}
\subsection{Spectral Reconstruction Time}
A typical use case for the PHOENIX generator may be to batch reconstruct spectra from an array of input stellar parameters. 
Therefore, there is some motivation to reduce the computational cost of this procedure to tractable levels. 
We evaluate the computational time needed to use the PHOENIX generator in two distinct ways. 
First, for a single input, which would be relevant in serial applications. 
Second, for an array of multiple inputs, which would be relevant in parallel applications.
The \texttt{RegularGridInterpolator} API allows for the passing in of an entire array of input coordinates to be evaluated at once. 
However using \texttt{blase} to reconstruct the spectrum from our interpolated state dictionary is always done serially, leading to what is actually more of a psuedo-parallel evaluation, but extremely performant nonetheless. 
Performance results are shown in \autoref{fig:figure5}, and we can see that the multi-reconstruction is much faster than a series of single reconstructions.

@@

\subsection{Inference Algorithm}
We elected to use Bayesian optimization as the inference algorithm, specifically the \texttt{gp\_minimize} function from the \texttt{scikit-optimize} library \citep{skopt}. 
This algorithm uses a Gaussian Process to model the objective function, which in this case was the RMS (Root-Mean-Square) loss between the interpolated spectrum $\mathsf{M}$ and the true spectrum $\mathsf{D}$, defined as:
\begin{gather}
    \mathcal{L} = \ev{(\mathsf{M} - \mathsf{D})^{\circ 2}}^{1/2}
\end{gather}
The optimizer was configured to first run 100 random evaluations to seed the surrogate model, then run 20 more evaluations now guided by the surrogate model. 
This totals to 120 evaluations, which was deemed sufficient for this study. 
One inference run takes on average just under 7.5 minutes to complete.

\subsection{Bayesian Optimizer Performance}
To test the performance of the inference algorithm, we used the PHOENIX subset itself. 
At first glance, this may seem circular, as the PHOENIX generator has memorized the PHOENIX subset, being able to reconstruct a PHOENIX spectrum when evaluated at that grid point. 
However, that stategy allows us to use the PHOENIX subset as test data in the context of Bayesian optimization.
The \texttt{gp\_minimize} surrogate model is seeded by random continuously sampled generator evaluations within the search space, \textit{not} grid points of the PHOENIX subset, meaning the surrogate model has no memorization to speak of. 
If the optimizer had been a grid-based strategy, this would not have been possible, because then the surrogate model would be affected by memorization.

@@

The results of the inference algorithm are as follows: $T_{\mathrm{eff}}$ was off from the true result by an average of 185 K or 2.6\%. 
$\log(g)$ was off by an average of 0.19 or 6.8\%. [Fe/H] was off by an average of 0.12 dex, which is 24\% of our search range.
From this, we can see that $T_{\mathrm{eff}}$ was the most accurately inferred parameter, followed by $\log(g)$, and then [Fe/H].

\section{Final Considerations}
The computer used for this study, Triton, has the following specifications:
\begin{table}[h!]
    \centering
    \begin{tabular}{ll}
        \hline
        CPU & AMD EPYC 7513\\
        RAM & 256 GB\\
        GPU & Nvidia A100 40GB ($\times$2)\\
        \hline
    \end{tabular}
    \caption{This machine was used for all computations, but not for generating visualizations. The EPYC 7513 is a 32c/64t CPU with a boost clock of 3.65 GHz. The A100 has 6912 CUDA cores.}
\end{table}

@@

@@

@@

    Future work would extend the reach of this study's algorithm to encompass other model grids such as the Sonora series of substellar models. 

@@

    Future work would involve constructing manifolds using more advanced methods, which would allow for much better generalization and lower disk utilization at the expense of some accuracy.

@@

    Future work would need to develop ways to tune these extrinsic parameters alongside stellar parameters.

@@

@@

@@

@@

In short, it is quite clear that \texttt{blase} is a long way from being a powerhouse in spectral inference, however it represents a step down a road not yet traveled, and the potential for future growth is immense. 
At the end of the day, we aim to develop a tool that can analyze spectra with the growing power of physics-informed machine learning.

>>>>

\begin{abstract}
    The advent of machine learning (ML) methods has and continues to revolutionize numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference.
    ML methods in spectroscopy are expected to perform with greater efficacy than conventional computational techniques.
    In this study we introduce a hybrid ML (HML) method that combines automatic differentiation (autodiff), interpolation, and Bayesian optimization to infer stellar parameters given stellar spectra.
    The stellar parameters we study are $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method has potential for extension to other properties such as [$\alpha$/Fe], C/O, and $f_{sed}$.
    We first use the semi-empirical approach to treating spectra as sets of tunable spectral lines introduced in \texttt{blase} \citep{blase}, contrasting with traditional data-driven approaches. 
    \texttt{blase} is used on 1,314 spectra from the PHOENIX synthetic spectral model grid \citep{PHOENIX}, identifying 128,723 unique spectral lines.
    For each of these lines, a manifold mapping stellar parameters to spectral line parameters is created using regular grid linear interpolation.
    These manifolds are aggregated to create the PHOENIX generator, sporting parallelized reconstruction of spectra at continuously valued points within the PHOENIX subset.
    Gaussian Process minimization is then used to infer stellar parameters by minimizing the root-mean-square (RMS) loss between the input spectrum and PHOENIX generator spectra.
    From testing, the inference error in $T_{\mathrm{eff}}$ was 185 K, $\log(g)$ was 0.19, and [Fe/H] was 0.12 dex.
    Our products are threefold: first an archive of the state dictionaries of the \texttt{blase} clones of the PHOENIX subset, second the PHOENIX generator itself, and third, the reconstruction and inference algorithm.
    This study represents a rudimentary proof of concept showing that semi-empirical HML is a viable, albeit fledgling alternative to traditional approaches.
\end{abstract}

\section{Introduction}
Stellar spectra are exceedingly rich sources of information about the stars that produce them. 
Spectra encode fundamental properties such as temperature, surface gravity, and chemical composition via their numerous absorption lines. 
Extrinsic properties such as the stellar radial velocity or the projected equatorial rotation shift the wavelengths of lines and broaden their widths, respectively. 
Observing the spectrum alters it again; the resolution, bandwidth, and other properties of the instrument change the fidelity at which we observe the spectrum, and can limit our ability to extract fundamental properties precisely.  
Observed stellar spectra thus represent extremely complex, data, influenced by multiple parameters, that have gone through multiple transformations before reaching our detectors.

The challenge lies in extracting as much information as possible from a stellar spectrum to get back its stellar parameters. 
Spectroscopic surveys such as APOGEE use their own in-house pipelines to extract stellar parameters from spectra (ASPCAP in the case of APOGEE), and these pipelines appear effective \citep{APOGEE, ASPCAP}. 
The core assumption for the vast majority of pipelines is that the data to be analyzed is a list of pixels. 
Analysis pipelines may be closed-source or limited to the scope of the survey itself, causing a sort of siloing effect among surveys. 
This motivates the development of more universal, instrument-agnostic open-source frameworks that can apply broadly to a range of spectral observations with relatively little tuning.

Multiple efforts have been made in this direction, treating spectra in different ways. 
The standard practice is to treat the wavelength and flux as simply two arrays and use bespoke algorithms tailored to a small number of well-calibrated spectral lines to obtain fundamental stellar parameters and chemical compositions \citep{pipeline1, pipeline2}. 
Other whole-spectrum fitting abstractions decompose model spectra into an eigenbasis, implementing the data-model comparison stage as a tractable regression, such as \texttt{starfish} \citep{starfish}.

\texttt{blase}, first presented in \citealt{blase}, takes an intermediate approach, treating spectra not as a set of pixels or a set of eigenbasis coefficients but as a set of spectral lines, specifically Voigt profiles.
Each of these approaches has tradeoffs, but one key computational advantage of  \texttt{blase} is that it can fit the nonlinear spectral line parameters with automatic differentiation (autodiff).
In a nutshell, autodiff is technology that tracks transformations made to data using the chain rule, even being able to differentiate control flow transformations involving if-else blocks and the like.
With the gradient obtained from autodiff's chain rule, we can optimize model parameters by just going against the direction of the gradient (because we usually want to minimize something such as a loss function, we go in the direction of greatest decrease, which is always the negative of the gradient).
We see autodiff used in other astrophysical software such as \texttt{exojax} and \texttt{wobble} \citep{exojax, wobble}, but the recasting of spectra into sets of inherently nonlinear spectral lines positions \texttt{blase} as a unique and promising semi-empirical tool.
The original \texttt{blase} paper demonstrated the ability to tune spectral lines with autodiff, but only on a few hand-picked fiducial models, not scaled out to an entire model grid.

Ideally this process could be monolithic, with the end-to-end spectral inference code powered by a single autodiffable machine learning framework, like PyTorch or JAX \citep{pytorch, jax}. 
However, here we separate the problem into three pieces. 
First, the scaling out of the blase method to 1,314 precomputed synthetic spectral model clones, yielding a downloadable archive of pretrained machine learning models with 128,723 unique spectral lines. 
Second, the creation of manifolds mapping stellar parameters to spectral line parameters using regular grid linear interpolation. 
Finally, we show how reconstructions of the spectra using said manifolds can be used for spectral inference.
An overview of this process is shown in \autoref{fig:figure1}.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure1}
    \caption{Overview of the process used in this study.}
    \label{fig:figure1}
\end{figure*}

@@

\section{Cloning the PHOENIX Model Grid}
\subsection{The PHOENIX Subset}
For the purposes of this study, we did not consider the full range of the PHOENIX synthetic spectral model grid. 
Instead, we focused on a subset of the grid, focusing on near solar metallicities and a broad range of effective temperature and surface gravity, with details given in \autoref{tab:table1}.

\begin{table}[h!]
    \centering
    \begin{tabular}{lll}
        \hline
        \bf{Parameter} & \bf{Symbol} & \bf{Interval}\\
        \hline
        Alpha Element Abundance & $\alpha$ & [0] dex\\
        Iron Abundance & [Fe/H] & [-0.5, 0] dex\\
        Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K\\
        Surface Gravity & $\log(g)$ & [2, 6]\\
        Wavelength & $\lambda$ & [8038, 12849] \AA\\
        \hline
    \end{tabular}
    \caption{The subset of the PHOENIX grid used in this study. 
    These limits were imposed to reduce the computational cost of the algorithms and to ensure a rectilinear parameter space in order to work with \texttt{scipy}'s \texttt{RegularGridInterpolator} \citep{scipy}. 
    The wavelength limits in particular roughly line up with that of the Habitable Zone Planet Finder (HPF) spectrograph \citep{HPF}. 
    This subset is comprised of 1,314 individual spectra: 73 unique $T_{\mathrm{eff}}$ values, 9 unique $\log(g)$ values, and 2 unique [Fe/H] values.}
    \label{tab:table1}
\end{table}

\subsection{Preprocessing with \texttt{gollum}}
First, the PHOENIX subset was programmatically retrieved with \texttt{gollum}, which downloaded them from the PHOENIX FTP server \citep{gollum}. The spectra were then put through a 
three-step preprocessing pipeline similar to that from \citealt{blase}.
\begin{enumerate}
    \item \textit{Blackbody Division}: Since the $T_{\mathrm{eff}}$ of each spectrum is known, the according blackbody spectrum was divided out.
    \item \textit{Percentile Normalization}: The spectra were normalized by dividing them by their 99th percentile in order to collapse the dynamic range of flux and only look at relative features.
    \item \textit{Continuum Normalization}: The spectra were further normalized by dividing them by a \nth{5} order polynomial continuum fit using a peak-finding filter in order to eliminate curvature that would inhibit line modeling.
\end{enumerate}
Mathematically, we can express the preprocessing as follows:
\begin{gather}
    \mathsf{\bar{S}} = \dfrac{\mathsf{S}}{\mathsf{B}\mathsf{Q}_5P_{99}}
\end{gather}
where $\mathsf{\bar{S}}$ is the preprocessed spectrum, $\mathsf{S}$ is the original spectrum,  $\mathsf{B}$ is the blackbody spectrum, $\mathsf{Q}_n$ is the $n^\mathrm{th}$ order polynomial continuum fit, and $P_n$ is the $n^\mathrm{th}$ percentile function. 
Arithmetic operations between arrays are assumed to be elementwise in all following notation.

\subsection{Line Identification with \texttt{blase}}
The next step was to convert the PHOENIX subset into a physically interpretable intermediate representation: a table of spectral line properties rather than a array of fluxes. 
We used \texttt{blase}, which models spectral lines as Voigt profiles and tunes the profiles to mimic the original PHOENIX spectrum with back propagation.
Back propagation, put simply, is the process of moving in the autodiff gradient field to update ML model parameters (like mentioned earlier, usually against the gradient because we minimize loss functions).
Four parameters were optimized: the line center $\mu$, the log-amplitude $\ln(a)$, the Gaussian width $\sigma$, and the Lorentzian width $\gamma$. 
The optimization used the Adam optimizer with a learning rate of 0.05 over 100 epochs \citep{adam}. 
In addition, we limited two custom parameters: wing cut to 6000 and prominence to 0.005. 
Wing cut is a parameter that determines the extent of the Voigt profile to evaluate, saving computational resources by not evaluating negligible line wings. 
Prominence sets a lower limit for the amplitude of detected lines, which also saves resources by disregarding shallow lines, so in our case we disregard lines that affect the spectrum by $<$ 0.5\%.
In short, smaller values for wing cut and smaller values for prominence both decrease the computational cost of \texttt{blase}'s cloning process at the expense of decreasing its accuracy slightly. 
We note that \texttt{blase} uses the pseudo-Voigt approximation, which saves on computational cost while remaining accurate to about 1\% \citep{pseudovoigt1, pseudovoigt2}. 
The pseudo-Voigt approximation uses a weighted average of a Gaussian and Lorentzian as opposed to a convolution.
\texttt{blase}'s pseudo-Voigt profile implementation uses the following:
\begin{subequations}
\begin{gather}
    \tilde{\mathsf{V}}_\mu(\lambda) = a\left[\eta \mathbf{L}(\lambda - \mu'; f) + (1 - \eta)\mathbf{G}(\lambda - \mu'; f)\right]\\
    \eta = \sum_{n=1}^3 \mathbf{u}_n\left(\dfrac{2\gamma}{f}\right)^n\\
    f = 32\sum_{n=0}^5 \mathbf{v}_n \left(\sqrt{2\ln(2)}\sigma\right)^{5-n}(\gamma)^n\\
    \nonumber \mathbf{u} = \mqty[1.36603 \\ -0.47719 \\ 0.11116] \quad \mathbf{v} = \mqty[1 \\ 2.69269 \\ 2.42843 \\ 4.47163 \\ 0.07842 \\ 1]
    \label{eq:equation2}
\end{gather}
\end{subequations}
where $\mathbf{L}$ and $\mathbf{G}$ are abbreviations for Lorentzian and Gaussian profiles, respectively. 
Notice that we use $\mu'$ instead of $\mu$ in the formula. 
This is because \texttt{blase} optionally allows the line center to shift slightly during optimization, and it is this shifted center which is used in computation. 
The individual Voigt profiles are still indexed by $\mu$ for cross-model line identification, explained in the next section. 
Once optimization was complete, the list of identified lines, was saved to a `state dictionary': a common representation for pre-trained machine learning models that can be stored to disk for reuse later. 
These are stored in the \texttt{.pt} file format for each of the 1,314 PHOENIX subset grid points. 
The total disk space these files take up is 465 MB (382 MB when downloaded as a zip archive). 
For reference, the storage space the PHOENIX subset takes up on disk is approximately 8.1 GB.
This represents a data compression factor of around 20 just by recasting the spectrum with \texttt{blase}.

@@

Spectral lines were often only detected in some spectra from the PHOENIX subset. 
In \autoref{fig:figure2}, we show that different grid points sport differing counts of detected spectral lines.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure2}
    \caption{Number of detected spectral lines at each grid point of 
    a slice of the PHOENIX subset at solar metallicity. We can see that
    the number of detected lines decreases with increasing $T_{\mathrm{eff}}$.
    Also note that from $T_{\mathrm{eff}} = 7000$ K onward, the spacing 
    between grid points increases from 100 K to 200 K.}
    \label{fig:figure2}
\end{figure*}

This missing-lines phenomenon can arise for different reasons.
Perhaps for astrophysical reasons: stellar atmospheres genuinely do not produce that line at detectable strength at the given temperature and surface gravity. 
Or alternatively, our line-finding and line-association algorithms missed it. 
Whatever the cause, these missing lines have immediate practical consequences. 
Rectilinear interpolation schemes break in regions where a line does not appear; you can't interpolate a quantity that simply doesn't exist.

@@

\section{Bayesian Inference and Testing}
\subsection{Spectral Reconstruction Time}
A typical use case for the PHOENIX generator may be to batch reconstruct spectra from an array of input stellar parameters. 
Therefore, there is some motivation to reduce the computational cost of this procedure to tractable levels. 
We evaluate the computational time needed to use the PHOENIX generator in two distinct ways. 
First, for a single input, which would be relevant in serial applications. 
Second, for an array of multiple inputs, which would be relevant in parallel applications.
The \texttt{RegularGridInterpolator} API allows for the passing in of an entire array of input coordinates to be evaluated at once. 
However using \texttt{blase} to reconstruct the spectrum from our interpolated state dictionary is always done serially, leading to what is actually more of a pseudo-parallel evaluation, but extremely performant nonetheless. 
Performance results are shown in \autoref{fig:figure5}, and we can see that the multi-reconstruction is much faster than a series of single reconstructions.

@@

\subsection{Inference Algorithm}
We elected to use Bayesian optimization as the inference algorithm, specifically the \texttt{gp\_minimize} function from the \texttt{scikit-optimize} library \citep{skopt}. 
This algorithm uses a Gaussian Process to model the objective function, which in this case was the RMS (Root-Mean-Square) loss between the interpolated spectrum $\mathsf{M}$ and the true spectrum $\mathsf{D}$, defined as:
\begin{gather}
    \mathcal{L} = \ev{(\mathsf{M} - \mathsf{D})^{\circ 2}}^{1/2}
\end{gather}
The optimizer was configured to first run 100 random evaluations to seed the surrogate model, then run 20 more evaluations now guided by the surrogate model. 
This totals to 120 evaluations, a large sample to create a fairly detailed surrogate model, then a moderately precise guided evaluation phase, which was deemed sufficient for this study. 
Fine-tuning these numbers is possible, but simply not warranted for a proof-of concept method.
One inference run takes on average just under 7.5 minutes to complete.

\subsection{Bayesian Optimizer Performance}
To test the performance of the inference algorithm, we used the PHOENIX subset itself. 
At first glance, this may seem circular, as the PHOENIX generator has memorized the PHOENIX subset, being able to reconstruct a PHOENIX spectrum when evaluated at that grid point. 
However, that strategy allows us to use the PHOENIX subset as test data in the context of Bayesian optimization.
The \texttt{gp\_minimize} surrogate model is seeded by random continuously sampled generator evaluations within the search space, \textit{not} grid points of the PHOENIX subset, meaning the surrogate model has no memorization to speak of. 
If the optimizer had been a grid-based strategy, this would not have been possible, because then the surrogate model would be affected by memorization.

@@

The results of the inference algorithm are as follows: $T_{\mathrm{eff}}$ differed from the true result by an average of 185 K or 2.6\%. 
$\log(g)$ differed by an average of 0.19 or 6.8\%. [Fe/H] differed by an average of 0.12 dex, which is 24\% of our search range.
From this, we can see that $T_{\mathrm{eff}}$ was the most accurately inferred parameter, followed by $\log(g)$, and then [Fe/H].

\section{Conclusions}
The computer used for this study, Triton, has the following specifications:
\begin{table}[h!]
    \centering
    \begin{tabular}{ll}
        \hline
        CPU & AMD EPYC 7513\\
        RAM & 256 GB\\
        GPU & Nvidia A100 40GB ($\times$2)\\
        \hline
    \end{tabular}
    \caption{This machine was used for all computations, but not for generating visualizations. The EPYC 7513 is a 32c/64t CPU with a boost clock of 3.65 GHz. The A100 has 6912 CUDA cores.}
\end{table}

@@

    Users would be able to fit a greater variety of stellar spectra in many different wavelength regimes.

@@

    This takes burden off the users to truncate their data to the PHOENIX generators, making use simpler.

@@

    Future work would extend the reach of this study's algorithm to encompass other model grids such as the Sonora series of substellar models, ATLAS \citep{atlas}, and coolTLUSTY \citep{coolTLUSTY}, reaching practitioners studying various types of stars and spectra.

@@

    Future work would involve constructing manifolds using more advanced methods, which would allow for much better generalization, high speed, and lower disk utilization at the expense of some accuracy.

@@

    Future work would need to develop ways to tune these extrinsic parameters alongside stellar parameters, enabling users to optimize these frequently-observed extrinsic parameters on top of the base stellar parameters.

@@

    This has the potential to greatly increase the speed of this algorithm, depending on how much overhead is avoided with a bespoke implementation.

@@

    Having more interpretability would let scientists actually study certain species and their spectral lines.

@@

    This would increase accuracy for end users.

@@

In this study, we have presented a proof-of-concept algorithm that interpolates a subset of the PHOENIX spectral model grid and then uses GP minimization for an inference algorithm.
We create state dictionaries for all spectra in the PHOENIX subset, lossily distilling the spectra with a ~20 data compression factor.
We create the PHOENIX generator, enabling anyone to create reconstructions of PHOENIX spectra with continuously valued stellar parameters.
We introduce and test our inference algorithm, with average absolute deviations from true values of 185 K in 
$T_\mathrm{eff}$, 0.19 in $\log(g)$, and 0.12 dex in [Fe/H].
The methods discussed here represent a step down a road not traveled in spectral inference, and have the potential to become more advanced in the future by fully utilizing the strengths of physics-informed machine learning.
